Propositions as Sessions, Mechanically

### 2015-04-09 paper/formalisation.tex
Working on the issues related to reduction relation definition.

Initially, an implicit notion of weakening was chosen so as to be in line with
\citeauthor{Wadler:2014}'s definitions. However, explicit weakening is
required to differentiate the principal cut elimination rules for the
interactions between servers and clients (spawning and weakening). The paper
presentation does not specify where weakening could occur; it is tacitly
assumed weakening can always be pushed further up the derivation tree to a
base rule e.g. \coqe$cp_fwd$ rule. Contraction is not supported in the
formalisation due to the difficulty in specifying its principal cut
elimination rule. In order to specify the rule one requires to show the
following:
\begin{gather*}
\tm{\nu A.(!0(A).P \mid ?z[[A]].Q)}
\quad \becomes \quad \\
\tm{\nu A.(!0(A).P \mid \nu A.(!0(A).P \mid \key{contract}_{\Gamma} Q))}
\end{gather*}

where $\tm{?z[[A]].Q}$ denotes contraction by $z$ of two channels of type $A$
in $Q$, $\tm{P^y} \tpvdash \tp{{?\Gamma} \comma \tmof{y}A}$ for some
sufficiently fresh name $y$ and $\key{contract}_{\Gamma} \tm{Q}$ contracts
$\tm{Q}$ with the names in $\Gamma$.

Unfortunately, the reduction relation does not have knowledge of the
environment, $\Gamma$, typing process $\tm{P}$ because this information is
only available after the compound cut is destructed, yielding a derivation for
$\tm{P}$. This fact does not impact upon expressing weakening because only the
names (not the types in the bindings) within the environment are required for
weakening.

In the empty choice rule there may only be a single binding in the context
(cf. an arbitrary context in \citeauthor{Wadler:2014}'s
presentation~\cite{Wadler:2014}). \todo{look into a fix for the empty choice
  rule} While this diverges with the presentation of linear logic it became
necessary to do this because otherwise there would be no relationship between
the free variables within a process and its environment. no there is a clear
iff relationship which permits reduction rules to apply such as i think
weakening but it should be any rule which takes some fvs of proc and infers
something about environments from that...otherwise what it is not at all clear
how to encode the reductions in a relation, the subprocesses environments are
created from destructing the environment

@@

% ... subject reduction ...

subject reduction proof proceeds by induction on the reduction relation then
is automatic by appeal to the proved well-typed reductions.

% ... cut elimination ....

@@

modularised development; tactics slow prover took a long time to reduce proofs
so split the structure of the development \todo{picture of development
  structure}.

@@

One of the design goals of the development was to re-use existing libraries
wherever possible. two such instances proved extremely useful. first is the
Metatheory library developed at upenn. second, also from upenn, is the library
of tactics from software foundations.

software foundations helped in many ways: simplifying tacticals, ease of
introducing proved assertions, etc. limited use of names introduced by prover
leading to more robust development.

used extensively: metatheory does not handle multiple binders (check this?)
maybe add a generic tactic to handle multiple binders... or perhaps add a
lemmas for handling binders

Linearity create issues with typing some examples; the splitting of the
environment was non-deterministic so required side-conditions to enforce a
certain splitting of binders. For example the rule for associativity of
cut. \todo{how to handle? is there a better way?}

>>>>

\subsubsection{Issues for Reduction}

A consequence of defining the reduction relation with respect to compound
process terms is that the subprocesses have implicit derivations which are
unavailable in the conclusion. For example, \coqe$reduce_spawn$ provides no
information about the environment which types \coqe$P$ in the statement of the
lemma. The derivation for \coqe$P$ (similarly, \coqe$Q$) is obtained by
inverting the hypothesis \key{WT} and specialising the resulting assumptions
to a newly introduced fresh variable (to replace the de Bruijn index). In the
case of \coqe$reduce_spawn$, two inversions are required to get derivations
for \coqe$P$ and \coqe$Q$. This property of the reduction relation affects the
definitions of the typing rules for CP, though it was not obvious at the time.
Notable examples are the structural rules, weakening and contraction, and
empty choice.

\textbf{Weakening and Contraction} Initially, an implicit notion of weakening
was chosen so as to be in line with \citeauthor{Wadler:2014}'s
definitions. However, explicit weakening is required to differentiate the
principal cut elimination rules for the interactions between servers and
clients; spawning and garbage collection (weakening). The paper presentation
does not specify where weakening could occur; it is tacitly assumed weakening
can always be pushed further up the derivation tree to a base rule
e.g. \coqe$cp_fwd$ rule.

\begin{samepage}
Contraction is not supported in the formalisation due to the difficulty in
specifying its principal cut elimination rule.  In order to specify the rule
one requires knowledge of the environment typing process \coqe$P$ with the
rule of the form:
\begin{gather*}
\tm{\nu A.(!0(A).P \mid ?z[[A]].Q)}
\quad \becomes \quad \\
\tm{\nu A.(!0(A).P \mid \nu A.(!0(A).P \mid \key{contract}_{\Gamma} Q))}
\end{gather*}

where $\Gamma$ is the assumed environment typing \coqe$P$, $\tm{?z[[A]].Q}$
denotes contraction by $z$ of two channels of type $A$ in $Q$ and
$\key{contract}_{\Gamma} \tm{Q}$ contracts $\tm{Q}$ with the bindings in
$\Gamma$. Unfortunately, as explained above, there is no way for the reduction
relation to obtain the environment $\Gamma$.
\end{samepage}

The reader may wonder how weakening is supported if not
contraction. Fortunately, weakening only requires the names (not the types in
the bindings) within the environment, and these may be obtained by relying on
the following lemma relating the free variables of a process and the domain of
its environment:
\begin{coq}
Lemma in_env_fv:
  forall P $\Gamma$ x (WT: P $\tpcp$ $\Gamma$),
    x $\in$ dom $\Gamma$ $\iff$ x $\in$ fv_proc P.
\end{coq}

\textbf{Empty Choice} The empty choice rule may only have a single binding in
the context (cf. an arbitrary context in \citeauthor{Wadler:2014}'s
presentation~\cite{Wadler:2014}). Adding an arbitrary environment to the empty
choice constructor would prohibit \coqe$in_env_fv$ since its process no longer
contains all the names of its environment. Not only would weakening be
inadmissible if \coqe$in_env_fv$ did not hold, but also axiom cut reduction:
\begin{coq}
Lemma reduce_axcut:
  forall P A (w : atom) $\Gamma$
         (NFV: w $\notin$ fv_proc P)
         (WT: $\nu$ A.(w $\link$ 0 $\mid$ P) $\tpcp$ $\Gamma$),
    (open_proc P w) $\tpcp$ $\Gamma$.
\end{coq}

The proof of \coqe$reduce_axcut$ needs to rename the variable chosen to open
the process \coqe$P$ in the conclusion. Initially, the variable is a fresh
\coqe$x$ introduced during the proof, but using the following
\coqe$typing_rename$ lemma, \coqe$w$ can replace \coqe$x$ by appeal to
\coqe$in_env_fv$:
\begin{coq}
Lemma typing_rename:
  forall $\Gamma$ P k (x y : atom) A
         (NINX: x $\notin$ dom $\Gamma$ $\cup$ fv_proc P)
         (NINY: y $\notin$ dom $\Gamma$ $\cup$ fv_proc P)
         (WTX: $\open{k}{x}{P}$ $\tpcp$ (x $\sim$ A) ++ $\Gamma$),
    $\open{k}{y}{P}$ $\tpcp$ (y $\sim$ A) ++ $\Gamma$.
\end{coq}

where $\open{k}{x}{P}$ denotes replacing de Bruijn index $k$ with free
variable $x$ in process $P$.

While this diverges with the linear logic rule for top, I believe it to be
sound within the CP/GV eco-system such that this variant of CP still leads to
a valid translation to GV. In a private communication, Philip Wadler echoed
these remarks but pointed out that this is likely more to do with GV
\todo{what's another way to say ``too small'' vs. ``too large''}.

@@

@@

\textbf{Modularised development} Not only is modular development an important
programming practice in general, but in the context of theorem proving it has
the advantage of reducing interactive load time. Executing proof scripts can
be time-consuming especially if those scripts make frequent use of decision
procedures performing proof search, then the execution time is proportional to
the size of the proof context and complexity of the goal. There were cases in
which the system crashed unexpectedly during interactive mode and it is
conceivable that having many large proofs in-memory could be a contributing
factor. For these reasons, the development is partitioned into GV and CP
sections each with sub-divided into definition, infrastructure and typing
modules (CP also has an additional module for reduction lemmas). The CPS
translation is also defined separately. Load time is considerably improved and
the consequences of a crash are mitigated.

\textbf{Representation} It is important not to go overboard in the
when developing not to use dependent types all the time without working out if
it will be to the advantage of the prover use dependent types and other power
with care for example we could don't want to use mutual inductive definitions
for the types construct since this is overly complicated for what we wish to
achieve and we don't want to index types with kinds since this complicates
further definitions it is about trying to use dependent types responsibly
without using them in an uncritical or thinking manner with great power comes
great responsibility

\textbf{Library reuse} One of the design goals of the development was to reuse
existing libraries wherever possible. Two such instances proved extremely
useful, first is the Metatheory library as already mentioned. Second, is the
library of tactics provided by the Software Foundations online book~\cite{SF}.
The tactics helped in many ways including simplifying introduction of
assertions, providing convenient shorthands and ... Traditionally, assertions
in Coq have to be explicitly written within a proof leading to the use of
empheremal names introduced by prover. If the proof of the assertion follows
from an already proved lemma and some assumptions in the context this is
unnecessary verbage and this is what the library allows to omit \todo{example
  of using the forwards tactic} an example of the so called forwards tactic is
given below the $\sim$ is a a shorthand for immediately applying \coqe$auto$
to all subgoals generated by the tactic.

\textbf{Linear environments} Linearity create issues with typing some
examples; the splitting of the environment was non-deterministic so required
side-conditions to enforce a certain splitting of binders. For example the
rule for associativity of cut. \todo{how to handle? is there a better way?}

@@
